---
title: "Staistic Modelling - Homework 1"
author: "Keerthi Gopalakrishnan(M12931398)"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Libraries needed

```{r , warning = FALSE}

library("readxl")
library("plyr")
library("gtools")
library("tidyr")
library("ggplot2")
library("purrr")
library("Hmisc")
library(kableExtra)
library(data.table)
library(MASS)

```

### Step 1 : Importing data from excel files

##### Code:

```{r , warning = FALSE}

FAA1 <- read_excel("/Users/keerthigopalakrishnan/Downloads/FAA1.xls")
FAA1$aircraft <- factor(FAA1$aircraft)
FAA2 <- read_excel("/Users/keerthigopalakrishnan/Downloads/FAA2.xls")
FAA2$aircraft <- factor(FAA2$aircraft)

```

\newline

### Step 2 : Structure of Datasets

##### Code:

```{r , warning = FALSE}

str(FAA1)
str(FAA2)

```

##### Observations:

1) There are 800 samples and 8 variables in FAA1. There are 150 samples and 7 variables in FAA2.
2) The two datasets are different:
* FAA1 contains an extra variable 'duration', which is not present in FAA2
* Sample size of FAA1 is much bigger
* FAA1 - 400 airbus and 400 boeing, FAA2 - 50 airbus and 100 boeing. Hence not equal proportions in each set

##### Conclusion:

There are differences in the two datasets.

### Step 3 : Merge

##### Code:

```{r , warning = FALSE}

FAA2$duration <- NA
FAA_Merged <- merge(FAA1, FAA2, all = T)
count(duplicated(FAA_Merged$speed_ground))
FAA_Merged_No_Dups <- FAA_Merged[!duplicated(FAA_Merged$speed_ground),]

```

##### Observations:

1) After merging the two datasets, 100 duplicates were observed in the resulting set. There are 850 unique rows
2) The duplicate rows are not of any use to the project and can be removed.

##### Conclusion:

There are 100 duplicate rows, they can be removed.

### Step 4 : Summary of Combined & No duplicate data set

##### Code:

```{r , warning = FALSE}

summary(FAA_Merged_No_Dups)

```

##### Observations:

1) Combined Data set - 850 observations and 8 variables.

##### Conclusion:

Summary statistics are as shown above.

### Step 5 : One PPT Summary of Findings

##### Observations:

1) There are totally 950 observations together from FAA1 and FAA2, 850 unique and 100 duplicate, final set contains 850
2) There are totally 692 missing values, where 50 are from duration and 642 from speed_air
3) The final combined dataset has 450 airbuses and 400 boeing aircrafts
4) Abnormal data was observed such as min distance = 34.08, max distance = 6533.05, Min duration = 14.76, min height = -3.546

### Step 6 : Abnormal values

##### Code:

```{r , warning = FALSE}
  
  FAA_clean <- subset( FAA_Merged_No_Dups, 
                                       (FAA_Merged_No_Dups$duration > 40 |
                                          is.na(FAA_Merged_No_Dups$duration))  &
                                       (FAA_Merged_No_Dups$speed_ground >= 30 ) &
                                     (FAA_Merged_No_Dups$speed_ground <= 140) &
                                       (FAA_Merged_No_Dups$height >= 6) &
                                       (FAA_Merged_No_Dups$distance < 6000) )

```

##### Observations:

* As per data dictionary there are 19 rows of abnormal values:
1) Min distance = 34.08 (not feasible), max distance = 6533.05 (which is more than runway length)
2) Min duration = 14.76 (not feasible)
3) min height = -3.546 (not feasible)
4) min speed_ground is less than 30mph & max is greater than 140

##### Conclusion:

There were 19 rows that contained abnormal values and have been removed.

### Step 7 : Data set with no duplicate and no abnormal values

##### Code:

```{r , warning = FALSE}
  
summary(FAA_clean)

```

##### Observations:

* There are 831 observations and 8 variables in the new dataset.
* The summary statistics is as shown above

##### Conclusion:

There are 831 observations and 8 variables in new dataset

### Step 8 : Histogram

##### Code:

```{r, warning = FALSE}

FAA_clean %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
    facet_wrap(~ key, scales = "free") +
    geom_histogram(bins = 10)

```

##### Observations:

* All the variables follow a distribution approximately normal, except speed_air and distance
* It can be said that distance & speed_air have right skewed histograms
* It has been observed that speed_air is missing for values that have speed_ground < 90
* Probably data containing the left half of the histogram observations have not been provided or was bounded

##### Conclusion:

From the right skewed histogram of distance & speed_air it can be said that values for speed_air is missing when speed_ground < 90 and the lower distance range data observations have not been provided

### Step 9 : PPT with 5 points

##### Observations:

* There are 831 observations and 8 variables in teh final cleaned dataset
* There are 678 missing values - 50 from duration and 628 from speed_air
* The distance is right skewed and so the analytics will apply to only this range of observations
* The speed_air is missing for speed_ground < 90
* There are 444 airbuses and 387 boeings

### Step 10 : Pairwise correlation

##### Code:

```{r, warning = FALSE}

FAA_clean$model <- ifelse(FAA_clean$aircraft == "airbus", 1, 0)

res2 <- rcorr(as.matrix(FAA_clean[,-1]) ,type = "pearson")
Corr_data <- res2$r

Variable = colnames(Corr_data[,-7])

c1 <- as.numeric(gsub("([0-9]+)*$", "\\1", Corr_data[7,-7]))

Sign_of_Correlation <- factor(c("negative",'negative','positive','positive','positive','positive','negative'))

table_1 <- data.frame( Variable = Variable,
           Size_of_Correlation = c1, Sign_of_Correlation = Sign_of_Correlation )

table_1 <- table_1[order(abs(table_1$Size_of_Correlation) ,  decreasing = TRUE),]

table_1

```


##### Observations:

* Distance has greatest correlation with speed_air, speed_ground
* Distance has relevant negative correlation with model type as well
* Distance has less correlation with the rest

##### Conclusion:

Distance has positive greatest correlation with speed_air, speed_ground of 0.94 and 0.86 respectively, and negative correlation with model of -0.23

### Step 11 : Pairwise correlation

##### Code:

```{r, warning = FALSE}

plot(FAA_clean$distance,FAA_clean$speed_air)
plot(FAA_clean$distance,FAA_clean$speed_ground)
ggplot(FAA_clean, aes(x = 1:831, y = distance, color = model)) + 
    geom_point(position = position_dodge(width = 0.4))
plot(FAA_clean$distance,FAA_clean$height)
plot(FAA_clean$distance,FAA_clean$pitch)
plot(FAA_clean$distance,FAA_clean$duration)
plot(FAA_clean$distance,FAA_clean$no_pasg)


```


##### Observations:

* Yes the correlation plots do align with the size of correlation factors shown above.
* speed in air and ground show a positive upward line of samples.
* the slight negativity can also be observed in duration and no_pasg

##### Conclusion:

Yes the plots align with the strength shown above.

### Step 12 : Airplane model


##### Conclusion:

Yes added the model make in 10 and 11

### Step 13 : Regressing Y/Distance

##### Code:

```{r, warning = FALSE}

m1 <- lm(distance ~ speed_air , data = FAA_clean)
m2 <- lm(distance ~ speed_ground , data = FAA_clean)
m3 <- lm(distance ~ duration , data = FAA_clean)
m4 <- lm(distance ~ pitch , data = FAA_clean)
m5 <- lm(distance ~ height , data = FAA_clean)
m6 <- lm(distance ~ model , data = FAA_clean)
m7 <- lm(distance ~ no_pasg , data = FAA_clean)


Variable = c("speed_air","speed_ground","duration","pitch","height","model","no_pasg")

c1 <- c(summary(m1)$coefficients[2,"Pr(>|t|)"] , summary(m2)$coefficients[2,"Pr(>|t|)"],
       summary(m3)$coefficients[2,"Pr(>|t|)"], summary(m4)$coefficients[2,"Pr(>|t|)"],
       summary(m5)$coefficients[2,"Pr(>|t|)"], summary(m6)$coefficients[2,"Pr(>|t|)"],
       summary(m7)$coefficients[2,"Pr(>|t|)"])

Sign_of_Coeff <- factor(c("positive",'positive','negative','positive','positive','negative','negative'))

table_2 <- data.frame( Variable = Variable,
           Size_of_P_value = c1, Sign_of_Coeff = Sign_of_Coeff )

table_2 <- table_2[order(abs(table_2$Size_of_P_value) ,  decreasing = FALSE),]

table_2

```


##### Observations:

* speed_ground, speed_air have insignificant p-values in comparison to 0.05, hence showing that they play a very relevant role in predicting distance.
* Model type, that is, airbus or boeing also has a relevance.
* The remaining parameters have questionable relevance, and may or may not make a big difference to the model


##### Conclusion:

speed in air and ground has found to be the most relevant variables in regressing

### Step 14 : Standardization

##### Code:

```{r, warning = FALSE}

# standardizing data
Faa_demo <- scale(FAA_clean[,-1])

m1 <- lm(distance ~ speed_air , data = as.data.frame(Faa_demo))
m2 <- lm(distance ~ speed_ground , data = as.data.frame(Faa_demo))
m3 <- lm(distance ~ duration , data = as.data.frame(Faa_demo))
m4 <- lm(distance ~ pitch , data = as.data.frame(Faa_demo))
m5 <- lm(distance ~ height , data = as.data.frame(Faa_demo))
m6 <- lm(distance ~ model , data = as.data.frame(Faa_demo))
m7 <- lm(distance ~ no_pasg , data = as.data.frame(Faa_demo))

Variable = c("speed_air","speed_ground","duration","pitch","height","model","no_pasg")

c1 <- c(summary(m1)$coefficients[2,"Pr(>|t|)"] , summary(m2)$coefficients[2,"Pr(>|t|)"],
       summary(m3)$coefficients[2,"Pr(>|t|)"], summary(m4)$coefficients[2,"Pr(>|t|)"],
       summary(m5)$coefficients[2,"Pr(>|t|)"], summary(m6)$coefficients[2,"Pr(>|t|)"],
       summary(m7)$coefficients[2,"Pr(>|t|)"])

Sign_of_Coeff <- factor(c("positive",'positive','negative','positive','positive','negative','negative'))

table_3 <- data.frame( Variable = Variable,
           Size_of_P_value = c1, Sign_of_Coeff = Sign_of_Coeff )

table_3 <- table_3[order(abs(table_3$Size_of_P_value) ,  decreasing = FALSE),]

table_3

```


##### Observations:

* After standardization,on doing a summary of the new standardized data it can be seen that the output is of a similar format.
* Though the regression results will be the same, standardizing the data will help match the data better.


##### Conclusion:

After standardization data is of a similar/consistent format as opposed to the original data set.

### Step 15 : Comparison of Tables

##### Code:

```{r, warning = FALSE}

Correlation_Value = table_1[-1,2]
Correlation_Sign = table_1[-1,3]
Regression_Value = table_2[-2,2]
Regression_Sign = table_2[-2,3]
Standardization_Regression_Value = table_3[-2,2]
Standardization_Regression_Sign = table_3[-2,3]

Variable = table_1[-1,1]

table_0 <- data.frame( Variable = Variable,
           Correlation_Value = Correlation_Value, Correlation_Sign = Correlation_Sign,
           Regression_Value = Regression_Value, Regression_Sign = Regression_Sign,
           Standardization_Regression_Value = Standardization_Regression_Value,
           Standardization_Regression_Sign = Standardization_Regression_Sign)

table_0

```


##### Observations:

* The three tables are consistent with each other.
* That is: The correlation table shows speed_air and speed_ground as most important, and so does the regression table. This is mainatined even after standardization.
* From the table, following the same order or relevance, and removing correlated elements, following is the most important variables in order : 
1) speed_ground
2) model
3) height
4) pitch
5) duration
6) no_pasg


##### Conclusion:

The three tables are consistent with each other, showing the same trend of variable relevance. Speed_air has been removed due to high correlation and low number of samples.

### Step 16 : Collinearity

##### Code:

```{r, warning = FALSE}

m1 <- lm( distance ~ speed_ground , FAA_clean)
m2 <- lm( distance ~ speed_air , FAA_clean)
m3 <- lm( distance ~ speed_ground + speed_air , FAA_clean)

coef(m1)
coef(m2)
coef(m3)

```


##### Observations:

1) Model 1 has a coefficent of 41.44 for speed_ground
2) Model 2 has a coefficent of 79.53 for speed_air
3) Model 3 has a coefficent of -14 for speed_ground and 93.96 for speed_air in speed_ground+speed_air
4) From these obervations, we can say that multicollinearity exists.
5) Speed_air and speed_ground are highly correlated with each other = 0.99 positive correlation, and hence any change in any one causes change in distance and on the second predictor variable. Hence a sign change in speed_ground is seen.
6) Speed_ground must be kept in model selection as we are predicting landing overrun, and speed_air has 600+ null values. Speed_ground and air are highly correlated and have more valuable samples which will result in a better prediction.


##### Conclusion:

Multicollinearity is observed because of high correlation between air_ground and air_speed. Hence speed_ground is retained in the model as that has a great positive influence on landing distance being overrun and more valuable samples.

### Step 17 : R-squared

##### Code:

```{r, warning = FALSE}

m1 <- lm( distance ~ speed_ground , data = as.data.frame(Faa_demo))
m2 <- lm( distance ~ speed_ground + model , data = as.data.frame(Faa_demo))
m3 <- lm( distance ~ speed_ground + model + height , data = as.data.frame(Faa_demo))
m4 <- lm( distance ~ speed_ground + model + height + pitch , data = as.data.frame(Faa_demo))
m5 <- lm( distance ~ speed_ground + model + height + pitch + duration , data = as.data.frame(Faa_demo))
m6 <- lm( distance ~ speed_ground + model + height + pitch + duration + no_pasg, data = as.data.frame(Faa_demo))

s1 <- summary(m1)
s2 <- summary(m2)
s3 <- summary(m3)
s4 <- summary(m4)
s5 <- summary(m5)
s6 <- summary(m6)


r_squared <- c( s1$r.squared , s2$r.squared , s3$r.squared, s4$r.squared, s5$r.squared,
                s6$r.squared)

p <- c (1 , 2, 3, 4, 5, 6)

plot(p,r_squared)


```


##### Observations:

1) From the plot it is clearly seen that model six has highest r-squared value of 0.8506
2) r-squared value always increases with number of parameters, and hence most of the times is not completely reliable for choosing model

##### Conclusion:

The sixth model is the better one out of the 6.

### Step 18 : Adjusted R-squared

##### Code:

```{r, warning = FALSE}

adjsuted_r <- c( s1$adj.r.squared , s2$adj.r.squared, s3$adj.r.squared, s4$adj.r.squared,
                s5$adj.r.squared, s6$adj.r.squared)

p <- c(1 , 2, 3, 4, 5, 6)

plot(p,adjsuted_r)


```


##### Observations:

1) Adjusted r-squared compensates for the added parameters. And hence provides a more reliable value.
2) From the plot it is clearly seen that model 5 has highest adjusted r-squared value of 0.849

##### Conclusion:

The 5 model is the better one out of the 6.

### Step 19 : AIC

##### Code:

```{r, warning = FALSE}

AIC_val <- c( AIC(m1) , AIC(m2), AIC(m3) , AIC(m4), AIC(m5) , AIC(m6) )

p <- c(1 , 2, 3, 4, 5, 6)

plot(p,AIC_val)


```


##### Observations:

1) AIC provides parameter compensation and keeps a check on overfitting.
2) The model with least AIC value is model 5

##### Conclusion:

Model 5 is better than the other 6 models

### Step 20 : Comparing Variable selection methods

##### Code:

```{r, warning = FALSE}

table_r <- data.frame( Model_Num = p , R_squared = r_squared )
table_adj_r <- data.frame( Model_Num = p , Adjusted_R_squared = adjsuted_r )
table_aic <- data.frame( Model_Num = p , AIV_Val = AIC_val )

table_r <- table_r[order(abs(table_r$R_squared) ,  decreasing = TRUE),]
table_adj_r <- table_adj_r[order(abs(table_adj_r$Adjusted_R_squared) ,  decreasing = TRUE),]
table_aic <- table_aic[order(abs(table_aic$AIV_Val) ,  decreasing = FALSE),]

table_r
table_adj_r
table_aic

```


##### Observations:

1) The higher the adjusted-R and R-squared values the better is the model, and the lower is the AIC the better is the model.
2) From above results Adjusted-rsquared,a nd AIC indicate that model 5 is the best model, r-squared has it as second best which is good enough. Adjusted-r square is more relaible than r-squared.
3) Model 5 contains speed_ground, model, height, 

##### Conclusion:

Model 5 with speed_ground, model, height, pitch, duration is the best model in predicting landing overrun

### Step 21 : Variable selection based on automated algorithm

##### Code:

```{r, warning = FALSE}

fit1 <- lm(distance ~ ., data = as.data.frame(Faa_demo[,-4]) )
fit2 <- lm(distance ~ 1, data = as.data.frame(Faa_demo[,-4]) )
stepAIC(fit2,direction="forward",scope=list(upper=fit1,lower=fit2))

```

##### Observations:

1) The stepAIC function has recommended model 3 to be the better model out of all the available combinations, as opposed to the model 5 suggestion using the other AIC method
2) It shows the least AIC value of -1563 for model 3 out of all models, where as step 19 showed an AIC value for model 3 in the 4th position
3) The sign difference is because of some additive value seen in the automatic algorithm. However this does not add significance to the model prediction.
4) The suggested coefficients are shown above. This important feature is not present in AIC function.

##### Conclusion:
The two AIC methods suggested different number of parameters needed for prediction. Automated AIC definitely penalizes for more parameters in a more efficient manner as compared to the first AIC method. StepAIC also suggests the coefficients for these params.Making it a more convenient option to use. However, due to missing values speed_air had to be deleted from the data set in order to use stepAIC. stepAIC does not work well when many values are missing in a set.

